---
title: Volcano
author: ~
date: '2020-12-13'
slug: A-first-post
categories: []
tags: []
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

Using the volcano data from tinytuesday I will investigate volcanic eruptions and the affect they have on the environment. I will use the volcano, eruptions and sulfur datasets in order to explore this. The volcano dataset contains the variables volcano_number and volcano_name, which can be used to identify the volcanoes, population_within_5_km, which is the number of people within 5 kilometers of the volcano, latitude and longitude, doubles which can show where the volcanoes are and primary_volcano_type, a factor with 26 levels which shows what types the volcanoes are. The sulfur dataset contains the variables neem, which is the sulfur detected in ng/g from melting ice cores in Greenland, wdc, which is the sulfur detected in ng/g from melting ice cores in Antarctica, and Year, which is the year in CE that the ice cores were formed.

```{r, message=FALSE, include=FALSE}
library(here)
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)
```

```{r}
dat1<-read.csv(here::here("content","post","2020-12-10-a-first-post","data","volcano.csv"))
dat2<-read.csv(here::here("content","post","2020-12-10-a-first-post","data","eruptions.csv"))
dat3<-read.csv(here::here("content","post","2020-12-10-a-first-post","data","events.csv"))
dat4<-read.csv(here::here("content","post","2020-12-10-a-first-post","data","sulfur.csv"))
```

# Question 1:
To what extent does a volcano eruption impact the atmosphere? To explore this I will attempt to determine the relation between major volcanic eruptions (defined here as any eruption with a vei of 4 or higher) and spikes in sulfur levels in both Greenland and Antarctica. I will do this by creating a line graph with sulfur concentrations and plotting the major volcanic eruptions as dots on the date they occurred.
```{r, message=FALSE}
filtered <- dat1 %>%
  
  left_join(dat2, by = c("volcano_number","volcano_name","longitude","latitude"))%>%
  
  filter(vei >= 4, start_year >= 500,start_year <= 706)
ggplot()+
  
  geom_line(data=dat4, aes(year, neem, color = "blue"))+
  
  geom_line(data=dat4, aes(year, wdc, color = "red"))+
  
  geom_point(data=filtered, aes(start_year,0))+
  
  scale_color_manual(name="Location of Ice Cores",labels=c("Greenland","Antarctica"),values=c("blue"="blue","red"="red"))+
  
  ylab("Sulfur (ng/g)")+
  
  xlab("Year")+
  
  ggtitle("Year vs Sulfur")
```
Most major volcanic eruptions seem to coincide with a large spike in sulfur concentration in either Greenland or Antarctica. However, not all major volcanic eruptions seem to cause a spike in sulfur levels, which could possibly be because these volcanoes are too far away and lack the power to affect Greenland and Antarctica. This suggests the impact most volcanic eruptions can have on the atmosphere is limited and significantly decreases as distance increases. However, within their range, volcanoes seem to have a large, although temporary affect on the atmosphere.

# Question 2:
Are there any areas where a large amount of people live that are currently at risk from volcano eruptions and how many people are living near it? To determine this, I will create a map of the various volcanoes around the world with populations of at least 1000 people within 5 kilometers of it and color them based on the total population within 5 kilometers.
```{r, message=FALSE}
dat1 %>%
  
  left_join(dat2, by = c("volcano_number","volcano_name","longitude","latitude"))%>%
  
  group_by(volcano_name,volcano_number)%>%
  
  filter(start_year >= 1800, start_year == max(start_year),population_within_5_km>=1000)%>%
  
  select(region,longitude,latitude,population_within_5_km)%>%
  
  ggplot(aes(longitude,latitude,color=population_within_5_km/1000000))+
  
  geom_point()+
  
  borders("world")+
  
  coord_quickmap()+
  
  ggtitle("Locations of Recently Erupted Volcanoes with nearby People")+
  
  theme(legend.position="bottom")+
  
  scale_color_gradient(name="Population within 5 km (Millions)",low="blue",high="red")
```
Most of the active volcanoes seem to be located on the coasts of continents. For one of these volcanoes in Mexico there are about 5 million people living within 5 kilometers of it which could be hazardous for the many people living there if it erupts again.

# Question 3:
What volcanoes have erupted the most often?
```{r, message=FALSE}
dat1 %>%
  
  left_join(dat2, by = c("volcano_number","volcano_name","longitude","latitude"))%>%
  
  group_by(volcano_number,volcano_name)%>%
  
  summarize(total = n())%>%
  
  ungroup()%>%
  
  slice_max(total, n = 10)
```


# Conclusion:

