---
title: Australia Fires
author: ~
date: '2020-12-13'
slug: australia-fires
categories: []
tags: []
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, include=FALSE}
library(here)
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)
knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

```{r, include=FALSE}
Rainfall<-read.csv(here::here("content","post","2020-12-13-australia-fires","data","rainfall.csv"))
Temperature<-read.csv(here::here("content","post","2020-12-13-australia-fires","data","temperature.csv"))
Fires<-read.csv(here::here("content","post","2020-12-13-australia-fires","data","MODIS_C6_Australia_and_New_Zealand_7d.csv"))
```
Using the Australia fires data from tinytuesday I will investigate the location and frequency of the fires in Australia. I will use the rainfall and MODIS datasets to explore this.

# Question 1:
How many of these fires happened near the cities? I will examine this by counting the number of instances of fires with longitude and latitude equal to the city coordinates rounded to 0 decimal places.
```{r}
CityCoord <- Rainfall %>%
  select(city_name,lat,long)%>%
  rename(longitude=long,latitude=lat)%>%
  mutate(longitude = round(longitude,digits=0),latitude=round(latitude,digits=0))%>%
  group_by(city_name)%>%
  summarize(latitude=mean(latitude),longitude=mean(longitude))%>%
  mutate(longitude = round(longitude,digits=0),latitude=round(latitude,digits=0))
Fires %>%
  mutate(longitude = round(longitude,digits=0),latitude=round(latitude,digits=0))%>%
  left_join(CityCoord,by = c("longitude","latitude"))%>%
  na.omit(city_name)%>%
  count(city_name)%>%
  ggplot(aes(city_name,n))+
  geom_bar(stat="identity")+
  ylab("Number of nearby fires")+
  xlab("City")+
  ggtitle("Number of Nearby Fires vs Cities")
```

For every city except three, there were no fires near them.

# Question 2:
Where have fires happened the most? To determine this, I will round the longitudes and latitudes of all fires to two decimal places and count the 10 areas with the highest number of fires.
```{r}
Fires %>%
  mutate(longitude = round(longitude,digits=2),latitude=round(latitude,digits=2))%>%
  group_by(longitude,latitude)%>%
  summarize(total = n())%>%
  ungroup()%>%
  slice_max(total, n = 10)%>%
  ggplot(aes(longitude,latitude))+
  borders("world")+
  coord_quickmap(xlim = 110:160,ylim = -50:-10)+
  geom_point()
```

It seems the areas with the highest number of fires are all near the eastern coast of Australia.

# Question 3:
How has the frequency of fires changed over the 7 days? To determine this, I will create a line graph counting every instance of a fire over the days.
```{r}
Fires %>%
  count(acq_date)%>%
  ggplot(aes(acq_date,n))+
  geom_line(group = 1)+
  ylab("Number of Fires")+
  xlab("Date")+
  ggtitle("Number of Fires over time")
```

The amount of fires is mostly consistent with a spike on the second and seventh day, but there's a very sharp drop in the number of fires on the eigth day.