---
title: Australia Fires
author: ~
date: '2020-12-13'
slug: australia-fires
categories: []
tags: []
---
```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, tidy = TRUE)
```

```{r, message=FALSE, include=FALSE}
library(here)
library(ggplot2)
library(tidyverse)
library(readxl)
library(dplyr)
library(patchwork)
```

Using the Australia fires data from tinytuesday I will investigate the location and frequency of the fires in Australia. I will use the rainfall and MODIS datasets to explore this. The variables in the rainfall data are city_name, which is the name of the city where rainfall was measured, lat and long, which are the coordinates of the city, year, month and day, which together are the date the rainfall was measured on. The MODIS data has the latitude and longitude variables which show the coordinates of the fire's center, acq_date which is the date when the fire started over the 8 days measured and confidence which is the percent chance that the point recorded is actually a fire.

```{r}
Rainfall<-read.csv(here::here("content","post","2020-12-13-australia-fires","data","rainfall.csv"))
Temperature<-read.csv(here::here("content","post","2020-12-13-australia-fires","data","temperature.csv"))
Fires<-read.csv(here::here("content","post","2020-12-13-australia-fires","data","MODIS_C6_Australia_and_New_Zealand_7d.csv"))
```

# Question 1:
How many of these fires happened near the cities, and have the number of fires changed in these cities changed over time? I will examine this by counting the number of instances of fires with longitude and latitude equal to the city coordinates rounded to 0 decimal places.
```{r, message=FALSE}
CityCoord <- Rainfall %>%
  
  select(city_name,lat,long)%>%
  
  rename(longitude=long,latitude=lat)%>%
  
  mutate(longitude = round(longitude,digits=0),latitude=round(latitude,digits=0))%>%
  
  group_by(city_name)%>%
  
  summarize(latitude=mean(latitude),longitude=mean(longitude))%>%
  
  mutate(longitude = round(longitude,digits=0),latitude=round(latitude,digits=0))

Fires %>%
  
  mutate(longitude = round(longitude,digits=0),latitude=round(latitude,digits=0))%>%
  
  left_join(CityCoord,by = c("longitude","latitude"))%>%
  
  na.omit(city_name)%>%
  
  count(city_name)%>%
  
  ggplot(aes(city_name,n))+
  
  geom_bar(stat="identity")+
  
  ylab("Number of nearby fires")+
  
  xlab("City")+
  
  ggtitle("Total Number of Nearby Fires vs Cities")

Fires %>%
  
  mutate(longitude = round(longitude,digits=0),latitude=round(latitude,digits=0))%>%
  
  left_join(CityCoord,by = c("longitude","latitude"))%>%
  
  group_by(acq_date)%>%
  
  na.omit(city_name)%>%
  
  count(city_name)%>%
  
  ggplot(aes(acq_date,n))+
  
  geom_bar(stat="identity")+
  
  ylab("Number of nearby fires")+
  
  xlab("City")+
  
  ggtitle("Number of Nearby Fires vs Cities over time")+
  
  facet_wrap(~ city_name)+
  
  theme(axis.text.x = element_text(angle = -75,hjust=-.1))
```

There were fires only near three out of the six cities recorded, and it seems Sydney has by far the greatest number of nearby fires, having almost double the number of Perth and Melbourne. Perth, however, is the city that has experienced the greatest number of fires in a single day with 7 fires, whereas Melbourne and Sydney only had 6 at most. In addition, that day in particular is also the one day where no fire has occurred near Sydney.


# Question 2:
Where have fires happened the most? To determine this, I will round the longitudes and latitudes of all fires to one decimal place and count the 20 areas with the highest number of fires. I will then map their coordinates along with their position relative to the cities (shown as black squares) listed in the CityCoord data created in question 1.
```{r, message=FALSE}
FireMap <- Fires %>%
  
  mutate(longitude = round(longitude,digits=1),latitude=round(latitude,digits=1))%>%
  
  group_by(longitude,latitude)%>%
  
  summarize(total = n())%>%
  
  ungroup()%>%
  
  slice_max(total, n = 20)%>%
  
  rename("TotalFires" = "total")
```  
In order to count the number of fire detections close to each other to determine concentration I rounded each of the fires' longitudes and latitudes from three to one decimal point, and for that particular pair of longitude and latitude one fire case is added. After that, the 20 pairs of longitude and latitude with the highest number of fires are recorded in FireMap.
```{r, message=FALSE}
ggplot()+
  
  borders("world")+
  
  coord_quickmap(xlim = 113:155,ylim = -38:-10)+
  
  geom_point(data=FireMap,aes(longitude,latitude,color=TotalFires))+
  
  guides(size = FALSE)+
  
  scale_color_gradient(name="Total Fires in the Area",low="blue",high="red")+
  
  geom_point(data = CityCoord, aes(x=longitude,y=latitude,shape=15))+
  
  scale_shape_identity()+
  
  ggtitle("Areas With the Heaviest Concentration of Fires")

Fires %>%
  
  mutate(longitude = round(longitude,digits=1),latitude=round(latitude,digits=1))%>%
  
  group_by(longitude,latitude)%>%
  
  summarize(total = n())%>%
  
  ungroup()%>%
  
  slice_max(total, n = 20)%>%
  
  rename("Total Fires In Area" = "total")
```

It seems the areas with the highest number of fires are all near the Southeastern coast of Australia. It should be taken into account, however, that total number of fires in an area does not necessarily indicate how damaged the area was by fire, as it doesn't take into account the intensity of the fires. As such, one area can be more damaged by fires than another even with a lower total number of fires as long as the intensity of each of its fires is large enough. Since the 21 areas with the largest concentrations of fire detections are also generally near where there are cities, human activity may be one of the causes behind the large concentration of fires in that area.

# Question 3:
How has the frequency of fires changed over the 8 days, and is there any difference between the frequency of fires during the day and during the night? To determine this, I will create a line graph counting every instance of a fire over the days and faceting it based on whether it happened during the day or during the night. One thing to note however is that, as indicated by the confidence variable ranging from 0% to 100%, it is possible that some of the fire detections by the satellite listed in the data are not actually fires, so I will also make graphs for only fire detections above 50% confidence and at 100% confidence to see if there's any variation between them.
```{r, message=FALSE}
levels(Fires$daynight) <- c("Day", "Night")

Fires %>%
  
  group_by(daynight)%>%
  
  count(acq_date)%>%
  
  ggplot(aes(acq_date,n))+
  
  geom_line(group = 1)+
  
  ylab("Number of Fires")+
  
  xlab("Date")+
  
  ggtitle("Number of Fires over time (All)")+
  
  facet_wrap(~ daynight)+
  
  theme(axis.text.x = element_text(angle = -45,hjust=-.1))

Fires %>%
  
  filter(confidence >= 50)%>%
  
  group_by(daynight)%>%
  
  count(acq_date)%>%
  
  ggplot(aes(acq_date,n))+
  
  geom_line(group = 1)+
  
  ylab("Number of Fires")+
  
  xlab("Date")+
  
  ggtitle("Number of Fires over time (50% and above Confidence)")+
  
  facet_wrap(~ daynight)+
  
  theme(axis.text.x = element_text(angle = -45,hjust=-.1))

Fires %>%
  
  filter(confidence >= 100)%>%
  
  group_by(daynight)%>%
  
  count(acq_date)%>%
  
  ggplot(aes(acq_date,n))+
  
  geom_line(group = 1)+
  
  ylab("Number of Fires")+
  
  xlab("Date")+
  
  ggtitle("Number of Fires over time (100% Confidence)")+
  
  facet_wrap(~ daynight)+
  
  theme(axis.text.x = element_text(angle = -45,hjust=-.1))
```

It seems amount of fires is mostly consistent with a spike on day two and day seven for nighttime and on the fifth and seventh day for daytime, but there's a very sharp drop in the number of fires for both daytime and nighttime on the eighth day. In addition, although different confidence levels change the total number of fires, the overall trend is mostly the same for the three levels chosen.

# Conclusion:

There seems to be a very high concentration of fires in Southeastern Australia. It is possible that this is due to human activity because this is also where alot of cities are located. However, the amount of fires may not be what they seem, since the fire detections with 100% confidence by the satellite are only about half the total number of fire detections. In conclusion, there should be a bigger focus on fire protection in southeastern australia.